---
title: "Built-in graphics"
author: "Charles Eliot"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Built-in graphics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../tests")
```
In this note we describe the graphing functions built into the **pacehrh** package.

## First things first ...

... creating some data.

```{r include=TRUE, eval=TRUE}
library(pacehrh)

pacehrh::InitializePopulation()
pacehrh::InitializeScenarios()
pacehrh::InitializeStochasticParameters()
pacehrh::InitializeSeasonality()
pacehrh::InitializeCadreRoles()

scenario <- "MergedModel"

pacehrh::SetGlobalStartEndYears(2025, 2050)

results <-
  pacehrh::RunExperiments(scenarioName = scenario,
                       trials = 25)
```
## Fertility and mortality rates

The function `PlotFertilityRatesStats()` creates graphs of the fertility and mortality rates used to generate population predictions.

* **type** : {"ribbon" | "lines" | "boxplot"}
* **se** : {TRUE | FALSE} - doesn't apply to `type = "boxplot"`
   + FALSE : confidence interval display based on standard deviation
   + TRUE : confidence interval display based on standard error
* **log** : {TRUE | FALSE}
   + FALSE : linear y-axis
   + TRUE : logarithmic (log10) y-axis

```{r include=TRUE, eval=TRUE, echo=TRUE, fig.height = 4, fig.width = 9}
g <- pacehrh::PlotFertilityRatesStats(results, type = "boxplot", log = FALSE)
print(g)

g <- pacehrh::PlotFertilityRatesStats(results, type = "boxplot", log = TRUE)
print(g)

g <- pacehrh::PlotFertilityRatesStats(results, se = FALSE, type = "lines")
print(g)

g <- pacehrh::PlotFertilityRatesStats(results, se = TRUE, type = "lines")
print(g)

g <- pacehrh::PlotFertilityRatesStats(results, se = FALSE, type = "ribbon")
print(g)

g <- pacehrh::PlotFertilityRatesStats(results, se = TRUE, type = "ribbon")
print(g)
```
