---
title: "The results structure"
author: "Charles Eliot"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The results structure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# The PACE-HRH model results structure

The script for a typical run of the PACE-HRH modeling engine looks like this: 

```{r include=TRUE, eval=FALSE}
library(pacehrh)

pacehrh::InitializePopulation()
pacehrh::InitializeScenarios()
pacehrh::InitializeStochasticParameters()
pacehrh::InitializeSeasonality()

scenario <- "ScenarioName"

results <-
  pacehrh::RunExperiments(scenarioName = scenario,
                       trials = 100)
```

In this note we describe the contents of the *results* data structure returned by ```RunExperiments()```.

## Structure

The *results* (plural) structure is an R list consisting of multiple *result* (singular)
structures, one per trial.

Each *result* structure consists of the following parts:

* **Clinical**: <TBD>
* **NonClinical**: <TBD>
* **NonClinicalAllocation**: <TBD>
* **NonProductive**: <TBD>
* **FTEs**: <TBD>
* **SeasonalityResults**: <TBD>
* **Population**: <TBD>
* **PopulationRates**: <TBD>







